Supplementary Material for \emph{A Guide to Pre-processing High-Frequency Animal Tracking Data}

### 2021-02-08 docs/ms_atlas_preproc_supplement.tex
new pdf and tex
add new tex and pdf

\usepackage{fontspec}
% use nice fonts if available else use boring defaults

\usepackage{lineno}
\KOMAoption{fontsize}{10pt}

\IfFontExistsTF{Times New Roman}{\setmainfont[]{Times New Roman}}{} 
\IfFontExistsTF{Arial}{\setsansfont[]{Arial}}{}
\IfFontExistsTF{Fira Code}{\setmonofont{Fira Code}}

@@

\title{Supplementary Material for \emph{A Guide to Pre-processing High-Frequency Animal Tracking Data}}
\author{Pratik R. Gupte \and Christine E. Beardsworth \and Orr Spiegel \and Emmanuel Lourie \and Sivan Toledo \and Ran Nathan \and Allert I. Bijleveld}
\date{2021-02-03}

@@

We began by visualising the data to check for location errors, and found a single outlier position approx. 15km away from the study area (Fig. 1.1, 1.2; main text Fig. 7.a).
This outlier was removed by filtering data by the X coordinate bounds using the function \texttt{atl\_filter\textbackslash{}\_bounds}; X coordinate bounds \(\leq\) 645,000 in the UTM 31N coordinate reference system were removed (n = 1; remaining positions = 50,815; Fig. 1.2).
We then calculated the incoming and outgoing speed, as well as the turning angle at each position using the functions \texttt{atl\_get\_speed} and \texttt{atl\_turning\_angle} respectively, as a precursor to targeting large-scale location errors in the form of point outliers.
We used the function \texttt{atl\_filter\_covariates} to remove positions with incoming and outgoing speeds \(\geq\) the speed threshold of 15 m/s (n = 13,491, 26.5\%; remaining positions = 37,324, 73.5\%; Fig. 1.3; main text Fig. 7.b).
This speed threshold was chosen as 5 m/s faster than the known boat speed, 10 m/s.
Finally, we targeted small-scale location errors by applying a median smooth with a moving window size \(K\) = 5 using the function \texttt{atl\_median\_smooth} (Fig. 1.4; main text Fig. 7.c).
Smoothing did not reduce the number of positions.
We thinned the data to a 30 second interval leaving 1,803 positions (4.8\% positions of the smoothed track)

@@

@@

>>>>

% \usepackage{fontspec}
% use nice fonts if available else use boring defaults

\usepackage{lineno}
% \KOMAoption{fontsize}{10pt}

% \IfFontExistsTF{Palatino}{\setmainfont[]{Palatino}}{} 
\usepackage{mathpazo}
\usepackage{helvet}
% \IfFontExistsTF{Arial}{\setsansfont[]{Arial}}{}
% \IfFontExistsTF{Fira Code}{\setmonofont{Fira Code}}

@@

\title{Supplementary Material for \emph{A Guide to Pre-processing High-Frequency Animal Tracking Data}}
\author{Pratik R. Gupte \and Christine E. Beardsworth \and Orr Spiegel \and Emmanuel Lourie \and Sivan Toledo \and Ran Nathan \and Allert I. Bijleveld}
\date{2021-02-08}

@@

We began by visualising the data to check for location errors, and found a single outlier position approx. 15km away from the study area (Fig. 1.1, 1.2; main text Fig. 7.a).
This outlier was removed by filtering data by the X coordinate bounds using the function \texttt{atl\_filter\_bounds}; X coordinate bounds \(\leq\) 645,000 in the UTM 31N coordinate reference system were removed (n = 1; remaining positions = 50,815; Fig. 1.2).
We then calculated the incoming and outgoing speed, as well as the turning angle at each position using the functions \texttt{atl\_get\_speed} and \texttt{atl\_turning\_angle} respectively, as a precursor to targeting large-scale location errors in the form of point outliers.
We used the function \texttt{atl\_filter\_covariates} to remove positions with incoming and outgoing speeds \(\geq\) the speed threshold of 15 m/s (n = 13,491, 26.5\%; remaining positions = 37,324, 73.5\%; Fig. 1.3; main text Fig. 7.b).
This speed threshold was chosen as 5 m/s faster than the known boat speed, 10 m/s.
Finally, we targeted small-scale location errors by applying a median smooth with a moving window size \(K\) = 5 using the function \texttt{atl\_median\_smooth} (Fig. 1.4; main text Fig. 7.c).
Smoothing did not reduce the number of positions.
We thinned the data to a 30 second interval leaving 1,803 positions (4.8\% positions of the smoothed track)

@@

\NormalTok{pal <-}\StringTok{ }\NormalTok{RColorBrewer}\OperatorTok{::}\KeywordTok{brewer.pal}\NormalTok{(}\DecValTok{5}\NormalTok{, }\StringTok{"Set1"}\NormalTok{)}
\NormalTok{pal[}\DecValTok{3}\NormalTok{] <-}\StringTok{ "seagreen"}
\NormalTok{pal[}\DecValTok{2}\NormalTok{] <-}\StringTok{ }\NormalTok{RColorBrewer}\OperatorTok{::}\KeywordTok{brewer.pal}\NormalTok{(}\DecValTok{5}\NormalTok{, }\StringTok{"Blues"}\NormalTok{)[}\DecValTok{5}\NormalTok{]}
\end{Highlighting}
\end{Shaded}

@@

\hypertarget{plot-figure-7}{%
\section{Plot figure 7}\label{plot-figure-7}}

Plotting code is not shown in PDF and HTML form, see the \texttt{.Rmd} file.
